---
title: "Homework 3"
author: "Mengyu Zhang"
date: "10/4/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 13,
  fig.asp = .6
  )
library(tidyverse)
library(viridis)
```

# Problem 1

### Description of the datase

```{r }
#load dataset
library(p8105.datasets)
data("instacart")

#primitive data structure
str(instacart)

```

The size of the data is `r nrow(instacart)` observations of `r ncol(instacart)` variables. The primitive data structures are integer and character, and the non-primitive data structure is `r mode(instacart)`. Key variables include the information of the orders like `r colnames(select(instacart, contains("order")))` and the information of users like `r colnames(select(instacart, contains("users")))` and the informaiton of products like `r colnames(select(instacart, contains("product")))`, `r colnames(select(instacart, contains("aisle")))`, `r colnames(select(instacart, contains("department")))`.

### Answer the questions
* How many aisles are there, and which aisles are the most items ordered from?

There are 134 aisles and "fresh vegetables" is the most items ordered from, and the number of order is 150609.
```{r}
#the number of aisles
knitr::kable(instacart %>%
  summarize(
  n_aisles = n_distinct(aisle)
  ))

#aisles that the most items ordered from
knitr::kable(instacart %>%
  group_by(aisle) %>%
  summarize(n_obs = n()) %>% 
  filter(min_rank(desc(n_obs)) == 1))
```

* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.


```{r scatterplot for the number of items ordered in each aisle}
# scatterplot for the number of items ordered in each aisle
instacart %>%
  group_by(aisle) %>%
  summarize(n_obs = n()) %>% 
  filter(n_obs > 10000) %>% 
  ggplot(aes(x = aisle, y = sort(n_obs))) + 
  geom_point(size = 2) + 
  labs(
    title = "The Number of Items Ordered in each Aisle",
    x = "Aisle Name",
    y = "The number of items"
  ) + 
  theme_bw() + 
  theme(axis.text.x = element_text(size = 9, vjust = 0.5, hjust = 0.5, angle = 45))
```

* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r table showing the three most popular items in each of the aisles}
three_pop_item = instacart %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>%
  summarize(n_obs = n()) %>% 
  ungroup() %>% 
  group_by(aisle) %>% 
  filter(min_rank(desc(n_obs)) < 4) 

knitr::kable(three_pop_item)
```

* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r table showing the mean hour of the day}
# table showing the mean hour of the day
mean_hod_dow = instacart %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hod = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hod
  )

knitr::kable(mean_hod_dow, digits = 3)
```


# Problem 2

### Import data
```{r, include=FALSE}
data("brfss_smart2010")
force(brfss_smart2010)
```

First, do some data cleaning:

* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r clean data}
#clean data
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health", 
         response == "Excellent" | 
           response == "Very good" | 
           response == "Good" | 
           response == "Fair" | 
           response == "Poor") %>% 
  mutate(
    response = factor(
      response, 
      levels = c("Poor", "Fair", "Good", "Very good", "Excellent")
      )
    )
```

Using this dataset, do or answer the following (commenting on the results of each):

* In 2002, which states were observed at 7 or more locations? What about in 2010?

In 2002 we can see there are 6 states were observed at 7 or more locations shown as table 1, and in 2010, there are 14 states were observed at 7 or more locations shown as table 2. 

**Comment:** the increasing number of observed locations shows that BRFSS has become a powerful tool for targeting and building health promotion activities. The benefits of this system have been brought to society, like addressing urgent and emerging health issue.

```{r popularity}
# in 2002
knitr::kable(
    brfss %>% 
      filter(year == 2002) %>% 
      group_by(locationabbr) %>% 
      distinct(locationdesc) %>% 
      count() %>% 
      filter(n >= 7),
    caption = "Table 1: In 2002, 6 states were observed at 7 or more locations"
  )

#in 2010
knitr::kable(
    brfss %>% 
      filter(year == 2010) %>% 
      group_by(locationabbr) %>% 
      distinct(locationdesc) %>% 
      count() %>% 
      filter(n >= 7),
    caption = "Table 2: In 2010, 14 states were observed at 7 or more locations"
  )
```


* Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

**Comment:** 

```{r Construct a dataset and Make a “spaghetti” plot}
# construct the dataset
mean_data_value = brfss %>% 
  filter(response == "Excellent") %>% 
  group_by(year, locationabbr) %>% 
  summarize(mean_data_value = mean(data_value, na.rm = TRUE)) %>% 
  rename(state = locationabbr) # change variable name

#knitr::kable(mean_data_value)

# “spaghetti” plot
mean_data_value %>% 
  ggplot(aes(x = year, y = mean_data_value)) +
  geom_line(aes(color = state)) + 
  scale_color_viridis(discrete=TRUE) + 
  labs(
    title = "Mean data value over time within a state",
    y = "Mean Data Value"
  ) + 
  theme_bw()
```

* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

**Comment:**

```{r distribution of data_value}
brfss%>% 
  filter(year == 2006 | year == 2010, 
         locationabbr == "NY") %>% 
  select(year, locationdesc, response, data_value) %>% 
  mutate(locationdesc = forcats::fct_reorder(locationdesc, year, median)) %>%
  ggplot(aes(x = response, y = data_value)) + 
  geom_point() + 
  facet_grid(year ~ locationdesc)
  
brfss%>% 
  filter(year == 2006 | year == 2010, 
         locationabbr == "NY") %>% 
  select(year, locationdesc, response, data_value) %>% 
  ggplot(aes(x = response, y = data_value)) + 
  geom_point(aes(color = locationdesc), size = 8, alpha=.6) + 
  facet_grid(.~year) + 
  scale_color_viridis(name = "Locations in NY state", discrete=TRUE)

```

# Problem 3
* Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r }
accel_data = read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  

```


* Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

* Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this grap




